---
title: "GENIE BigQuery Example"
output:
  html_document:
    df_print: paged
---

## Load bigrquery package

You will have to run `bg_auth()` in your console to connect to bigquery after loading the bigrquery package.

```{r, results='hide'}
library(bigrquery)
# bq_auth()  # This line will have to be run non-interactively in your console
billing = "project-genie-query-prod"
```

## GENIE version and tables

Retrieve consortium release version in BigQuery.

```{r, echo=T, results='hide'}
metadata <- bq_dataset_meta("project-genie-query-prod.consortium", fields = NULL)
metadata$labels$release
```
Current version: `r metadata$labels$release`

Retrieve available tables
```{r, echo=T, results='hide'}
tables <- bq_dataset_tables("project-genie-query-prod.consortium")
table_names = lapply(tables, function(x) {
  paste(x$project, x$dataset, x$table, sep=".")
})
table_names
```

* `r table_names[[1]]`
* `r table_names[[2]]`
* `r table_names[[3]]`
* `r table_names[[4]]`


## Simple Query

This query retrieves all the patients that have a KRAS mutation

```{r, results='hide'}
sql <- paste(
  "SELECT sample.SAMPLE_ID, mutation.Hugo_Symbol,",
  "       mutation.Start_Position, sample.PATIENT_ID",
  "FROM (select * from project-genie-query-prod.consortium.mutation ",
  "      where Hugo_Symbol = 'KRAS') as mutation",
  "JOIN project-genie-query-prod.consortium.sample ",
  "ON sample.SAMPLE_ID = mutation.Tumor_Sample_Barcode"
)
tb <- bq_project_query(billing, sql)
kras_patients = bq_table_download(tb)
```

```{r}
knitr::kable(kras_patients[1:5, ])
```

# Advanced query

```{r, results='hide'}
sql = paste(
  "WITH genes AS (",
  "  SELECT * FROM UNNEST([",
  "       'ABL1',   'AKT1',    'ALK',    'APC',",
  "       'BRAF',   'CDH1',    'CDKN2A', 'CSF1R',",
  "       'CTNNB1', 'EGFR',    'ERBB2',  'ERBB3',",
  "       'ESR1',   'FBXW7',   'FGFR1',  'FGFR2',",
  "       'FGFR3',  'FLT3',    'GNA11',  'GNAQ',",
  "       'GNAS',   'HRAS',    'IDH1',   'IDH2',",
  "       'JAK2',   'JAK3',    'KIT',    'KLLN',",
  "       'KRAS',   'MAP2K1',  'MET',    'MLH1',",
  "       'MPL',    'MYC',     'NOTCH1', 'NRAS',",
  "       'PDGFRA', 'PIK3CA',  'PIK3R1', 'PTEN',",
  "       'PTPN11', 'RB1',     'RET',    'RUNX1',",
  "       'SMAD4',  'SMARCB1', 'SRC',    'STK11',",
  "       'TP53',   'VHL',     'WRAP53'",
  "   ]) AS symbol",
  "), luad AS (",
  "  SELECT DISTINCT patient_id, sample_id",
  "    FROM `project-genie-query-prod.consortium.sample` ",
  "   WHERE cancer_type_detailed = 'Lung Adenocarcinoma'",
  "), patient AS (",
  "  SELECT COUNT(DISTINCT patient_id) total",
  "    FROM `project-genie-query-prod.consortium.mutation` m, luad",
  "   WHERE m.Tumor_Sample_Barcode = luad.sample_id",
  ")",
  "  SELECT m.Hugo_Symbol, ",
  "         COUNT(DISTINCT luad.patient_id)/(SELECT total FROM patient) mut_freq",
  "    FROM `project-genie-query-prod.consortium.mutation` m, luad",
  "   WHERE m.Hugo_Symbol IN (SELECT symbol FROM genes)",
  "     AND m.Variant_Type ='SNP'",
  "     AND m.Tumor_Sample_Barcode = luad.SAMPLE_ID ",
  "GROUP BY m.Hugo_Symbol ",
  "ORDER BY mut_freq desc"
)
tb <- bq_project_query(billing, sql)
mutation_freq = bq_table_download(tb)

```

```{r}
knitr::kable(mutation_freq)
```





